# A very statistical day


Lets begin with some matrices. Matrices are a good way to express a system of linear equations in a more compact form. It's benefits include helping the reader to manipulate a system of linear equations easily and spot insights easily. 


- $\begin{array}{ccc}
{\sigma}^{2} & 1\\
1 & {\sigma}^{2}
\end{array}$

$\small
X = \begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix}  + \begin{bmatrix}1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\\end{bmatrix} + \begin{bmatrix} 1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\\end{bmatrix}$

- **Note**
    - This is a matrix
    - This is the first point
        - This is the seccond point
   
        
        
$y = mx + c + d + e + f + g + h + i + j + k + l + m + n + o + p$

```{r }
library(ggplot2)

carData::Bfox


hist_plot <- ggplot(data = carData::Bfox, aes(x = debt)) + geom_histogram()
```

```{r echo = F}
hist_plot
```


