---
title: "Statistics"
author: "Samuel Mak"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::bs4_book
description: |
  This is a minimal example of using the bookdown package to write a book.
  The output format for this example is bookdown::bs4_book.
  
  
---

# The Linear Regression Model


**The Linear Model**

- **Description**
    - A model that models/describes the relationship between variables with the linear function
- **Mathematics** 
    - $y = mx + c$
    
---    
**Regression**

- **Concept**
    - The process of estimating the relationship between variables

------   
    
**Linear Regression**

- **Concept**
  - The process of estimating the relationship between variables with a linear model by fitting a straight line to the data
  
---
**Curvilinear Regression**

  - **Concept**
    - The process of estimating the relationship between variables with curvilinear models by fitting a curve to the data



<!--chapter:end:index.Rmd-->

# The pool of tears


<!--chapter:end:02-tears.Rmd-->

# A caucus-race and a long tale

This is chapter 3, and I will talk about robust methods in this chapter. Robust statistics is a huge topic and it really needs a whole library to contain all of the knowledge in this topic. 


In this chapter, we will look at 

- Guitars
- Music
- Statistics


<p style="font-size:15px; color: black; text-align:left; line-height: 2; margin-left: -20; margin-right: -20;"> 
</p>

<!--chapter:end:03-race.Rmd-->

# A very statistical day


Lets begin with some matrices. Matrices are a good way to express a system of linear equations in a more compact form. It's benefits include helping the reader to manipulate a system of linear equations easily and spot insights easily. 


- $\begin{array}{ccc}
{\sigma}^{2} & 1\\
1 & {\sigma}^{2}
\end{array}$

$\small
X = \begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix} + 
\begin{bmatrix}
1 & 2\\
1 & 2\\
1 & 2
\end{bmatrix}  + \begin{bmatrix}1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\\end{bmatrix} + \begin{bmatrix} 1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\1 & 2 & 3 & 4 & 5 \\\end{bmatrix}$

- **Note**
    - This is a matrix
    - This is the first point
        - This is the seccond point
   
        
        
$y = mx + c + d + e + f + g + h + i + j + k + l + m + n + o + p$

```{r }
library(ggplot2)

carData::Bfox


hist_plot <- ggplot(data = carData::Bfox, aes(x = debt)) + geom_histogram()
```

```{r echo = F}
hist_plot
```



<!--chapter:end:04-Linear_Regression.Rmd-->

